---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gdslStats

<!-- badges: start -->
<!-- badges: end -->

The goal of gdslStats is to provide a set of tools to simplify the process of conducting the types of analyses normally covered in an introductory statistics course - data description, tabulation, correlation and OLS and logistic regression. Highlights of the package include the ability to create publication-standard of 1-, 2- and 3-way versions of cross-tabulations.

## Installation

You can install the development version of gdslStats from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("DrPaulWilliamson/gdslStats")
```

## Rounding

The `round` function provided as part of `gdslStats` is designed to to replace the `round` function supplied as part of R's `base` package. The `gdslStats` version of this function provides better support for the rounding of data objects that include columns (tables, matrices, data.frames and tibbles). Specifically, columns stored as factors or strings are ignored, rather than causing an error, and the option is provided to apply different roundings to different columns. In addition, for tables and tibbles, the way the data object is displayed is adjusted to match the rounding applied. Finally, more informative error messages are provided if asked to round the wrong kind of data.

```{r example_round}
library(gdslStats)
round( data_in = 1.1234 )

df <- data.frame( var = c("A", "B", "C"), v1 = rep(1.1234, 3), v2 = rep(1.1234, 3) )

round( df, dp = 1 )

round( df, dp = c(NA, 1, 2) )

#' # Example of table with as_string = FALSE to follow

round( tibble::as_tibble( df), dp = c(NA, 2, 1) )
```

## (Cross-)tabulation

The `tab` function provides a wrapper to various base r tabulation functions designed to simplify the process of tabulating survey or admin style datasets. The function supports the creation of one-, two- and three-way tabulations using counts or rates (joint, row or column proportions or percentages). The resulting table can optionally include row and/or column totals, whilst tables involving row or column rates can also include the base count (frequency) underpinning the rate. For three-way tables a further option is to include a final sub-table aggregating results across the third dimension. 

The output from the `tab` function is a data object of class `tab`, which is compatible with functions requiring `table` and `xtable` class inputs, but also compatible with the `publish_tab` fuction which can be used to convert any `tab` data object into a publication-standard table in one line of code.

```{r example_tab}
#' # Load toy dataset supplied with package
#' #load(survey)
#'
#' # Tabulation examples
#'
#' # Frequency table (counts), with and without missing values
#' survey |> tab( ~WorkStatus, na.rm = TRUE )
#' survey |> tab( ~WorkStatus, na.rm = FALSE )
#'
#' # Frequency table (proportions and cumulative proportions)
#' survey |> tab( ~ Age, measure = "col_prop" )
#' survey |> tab( ~ Age, measure = "col_prop", cum_sum = TRUE )
#'
#' # Two-way cross-tabulation (column percentages; row percentages with row totals)
#' survey |> tab( Sex ~ Age, measure = "col_pct" )
#' survey |> tab( Sex ~ Age, measure = "row_pct", totals = "row_col" )
#'
#' # Two-way cross-tabulation (row percentages with row totals and base count)
#' survey |> tab( Sex ~ Age, measure = "row_pct", totals = "col",
#'                base_count = TRUE )
#'
#' # Three-way cross-tabulation including counts aggregated across third dimension
#' survey |> tab( Health ~ Age + Sex, measure = "count", total_3d = TRUE )
#'
#' # All-singing all-dancing three-way table
#' survey |> tab( Health ~ Age + Sex, measure = "row_pct", totals = "row",
#'                base_count = TRUE, total_3d = TRUE )
#'
#' # Publication-standard version of a table
#' # (note: any rounding must take place prior to call to `publish_tab` )
#' survey |>
#'   tab( Age ~ Sex, measure = "row_pct", 
#'        totals = "row", base_count = TRUE ) |>
#'   round( dp = 0 ) |>
#'   publish_tab( )
```

## Grouping data

The `group_data` function groups continuous data into categories, providing user-friendly labels for the resulting grouped categories. It is a wrapper to the base R `cut( )` function.

```{r}
#' # Group original data into five equal-interval categories
#' group_data( data = survey$Height, breaks = 5 ) |> head( )
#'
#' # Compare original to grouped values
#' data.frame( original = survey$Height,
#'             grouped = survey |> dplyr::select( Height ) |> group_data( 5 )
#'           ) |> head( )
#'
#' # Class boundaries specified to n decimal places
#' group_data( data = survey$Height, breaks = 5, output.dp = 4 ) |> head( )
#'
#' # Report class mid-point instead of upper and lower bounds
#' group_data( data = survey$Height, breaks = 5, mid.point = TRUE ) |> head( )
#'
#' # Threshold for treating a data value as integer rather than continuous
#' group_data( data = survey$Height, breaks = 5, integer.dp = 0 ) |> head( )
#'
#' # Exclude data values equal to lower class boundary; include data values
#' # equal to upper class boundary
#' group_data( data = survey$Height, breaks = 5,
#'             include.lowest = FALSE, output.dp = 4 ) |> head( )
```
